AIS

### 2023-08-16 02_AIS.Rmd
overwrote old code, added Taiki's code

---
title: "AIS"
author: "Marina Bozinovic"
date: "2023-04-11"
output: html_document
---
```{r}
#Required Packages

library(tidyverse)
library(sf)
library(gganimate)
#install.packages("remotes")
#remotes::install_github("skgrange/gissr")
#devtools::install_github("skgrange/gissr")
library(gissr)

# Function
# Speeding up st_intersection (from https://github.com/r-spatial/sf/issues/801)

st_intersection_faster <- function(x,y,...){
#faster replacement for st_intersection(x, y,...)

  y_subset <-
    st_intersects(x, y) %>%
    unlist() %>%
    unique() %>%
    sort() %>%
    {y[.,]}

  st_intersection(x, y_subset,...)
}

```

```{r}
#########
# AIS
#########

# Need to add more AIS dates in data folder
AISlist <- list.files(path = paste0("data-raw/"), pattern = "AIS_", recursive = TRUE)

for (v in AISlist) {
  Adata <- read_csv(paste0("data-raw/", v), show_col_types = FALSE) %>%
    filter(Length >= 100) %>%                  # Filter out vessels over 100m in length
    dplyr::select(-SOG,-COG, -CallSign, -IMO, -Cargo, -Draft, -Status) %>%  #Remove unnecessary columns
    mutate(variable = paste0(substr(v, 16, 17), substr(v, 6, 7)))

  assign(paste0("AIS_", substr(v, 5, 10)), Adata)
}

AIS_sf <- AIS_091218 %>%
  st_as_sf(coords = c("LON", "LAT"), crs=4326, remove=F)
```

```{r}
##################################
# Create 10 km buffer around buoys (taken from M. McKenna's code)
# Should be malleable for each buoy's number to easily plug in.
##################################
tracks <- readRDS("data-raw/tracks.rda")

trk8 <- tracks %>% 
  dplyr::filter(station == "8") %>%
  st_as_sf(coords= c("Longitude", "Latitude"), crs=4326, remove=F)

pois <- st_transform(trk8, crs = 4326)    # Specify a crs
pts_buff <- st_buffer(pois, dist = 10000)  # Creates 10 km buffer around buoy points 10 km (10000m)
buff <- st_transform(pts_buff, crs = 4326)  # Specify crs for new buffer

### Crop full AIS to 10km buffer around drift 8 ###
AIS_crop <- st_crop(x = AIS_sf, y = buff)
```

```{r}
# Intersect buffer with AIS. Starting this on first 10 rows to see how slowly it goes. 
## FIX ME
#inters <- st_intersection(buff, AIS_crop[1:140,])

#Another way to view AIS points that fall within buffer. DOES NOT consider time.
AIS_within_buff <- AIS_crop[buff,]
f <- buff[AIS_within_buff,]
by <- join_by(closest(UTC >= BaseDateTime))

d <- full_join(buff, as.data.frame(AIS_within_buff), by)

e <- d[!duplicated(d[c('BaseDateTime')]),]

trk_AIS <- left_join(as.data.frame(trk8), as.data.frame(AIS_within_buff), by)

proj4string <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
trk_AIS_sf <- trk_AIS %>% st_as_sf(coords = c("Longitude", "Latitude"), crs=projcrs, remove=F)# %>%
   st_set_crs(4326)
trk_AIS_sf <- st_transform(trk_AIS_sf, crs = 4326)

#inters <- st_intersection_faster(buff, AIS_crop[1:140,])
#inters2 <- st_intersection_faster(buff, AIS_crop[141:200,])
#inters3 <- st_intersection_faster(buff, AIS_crop[201:300,])

#Another way to view AIS points that fall within buffer. DOES NOT consider time.

st_is_within_distance(AIS_crop, pois, dist = 10000)
st_distance(AIS_crop, pois, dist_fun = 10000)

st_set_geometry(AIS_crop, value = NULL)
st_set_geometry(pois, value = NULL)

AIS_crop <- as_Spatial(AIS_crop)
pois <- as_Spatial(pois)
AIS_within_buff <- as_Spatial(AIS_within_buff)
z <- sp_distance(AIS_crop, pois, features = TRUE, unit = "km")

data_places <- pois@data
data_places$distance <- sp_distance(AIS_within_buff, pois, features = TRUE, unit = "km")
View(data_places)

rgeos::gWithinDistance(AIS_crop, pois, features = TRUE, dist = 10000)
```
```{r}
# For each row in AIS df,
  # calculate distance to closest drift
  # return smallest distance
  # attach column to df

# 
(geosphere::distHaversine(AIS_within_buff[9:20,], pois))/1000
```

```{r}
# Plot showing 10km buffer around drift 8
#tmap_mode("plot")
#tm_shape(buff) + tm_symbols(col = "blue", border.col = NA) +
#  tm_shape(trk) + tm_dots(col = "red")
```

plots
```{r}
# Plot
ggplot(tracks_SS_08, aes(x = long, y = lat.x)) + geom_point() 
 + geom_sf(data = bbox8, aes(col = "red"))

# Plot showing AIS bounding box around drift 8 tracks
ggplot() + geom_sf(data = trk_AIS_sf, color = "red")
ggplot() + geom_sf(data = buff[lgth,] ,color = "red")

ggplot() +
  geom_sf(data = calif) + coord_sf(default_crs = sf::st_crs(4326)) + theme_bw()
 # geom_sf(data = AIS_sf)

p <- ggplot() + geom_sf(data = buff, mapping = aes(x=Longitude,y=Latitude)) +
  geom_point(data = AIS_within_buff, mapping = aes(x=LON,y=LAT, group = VesselName), color = "red")

anim <- p +
  transition_reveal(along = BaseDateTime)+
  ease_aes('linear')+
  #shadow_wake(0.1) +
  ggtitle("Date: {frame_along}")

animate(anim, nframes = 180, fps = 3)
```

>>>>

---
title: "AIS"
author: "Marina Bozinovic"
date: "2023-04-11"
output: html_document
---

#Introduction
AIS download from Marine Cadastre and match to buoy datasets.

Required Packages
```{r, }
library(PAMscapes)
library(tidyverse)
```

-----------------------
From Taiki's AIS Workflow

```{r}
#marcadFiles07 <- downloadMarCadAIS(tracks_SS_07, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles08 <- downloadMarCadAIS(tracks_SS_08, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles10 <- downloadMarCadAIS(tracks_SS_10, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles12 <- downloadMarCadAIS(tracks_SS_12, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles13 <- downloadMarCadAIS(tracks_SS_13, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles14 <- downloadMarCadAIS(tracks_SS_14, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles16 <- downloadMarCadAIS(tracks_SS_16, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles18 <- downloadMarCadAIS(tracks_SS_18, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles19 <- downloadMarCadAIS(tracks_SS_19, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles20 <- downloadMarCadAIS(tracks_SS_20, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles21 <- downloadMarCadAIS(tracks_SS_21, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles22 <- downloadMarCadAIS(tracks_SS_22, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles23 <- downloadMarCadAIS(tracks_SS_23, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
```

Subset to US West Coast
```{r}
subsetMarCadAIS(inDir="D:/AIS", outDir="D:/AIS/AIS_West", name='West_', 
                latRange=c(20,50), lonRange=c(-140,-110),
                overwrite=FALSE, progress=TRUE)
```
Now read into R only ais points that come in contact with buoys
```{r}
ais_near_07 <- readLocalAIS(tracks_SS_07, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
   filter(vesselLength >= 100)  #remove vessels under 100m in length

ais_near_08 <- readLocalAIS(tracks_SS_08, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
   filter(vesselLength >= 100)  #remove vessels under 100m in length

ais_near_10 <- readLocalAIS(tracks_SS_10, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
   filter(vesselLength >= 100)  #remove vessels under 100m in length

ais_near_12 <- readLocalAIS(tracks_SS_12, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
   filter(vesselLength >= 100)  #remove vessels under 100m in length

ais_near_13 <- readLocalAIS(tracks_SS_13, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
   filter(vesselLength >= 100)  #remove vessels under 100m in length

ais_near_14 <- readLocalAIS(tracks_SS_14, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
   filter(vesselLength >= 100)  #remove vessels under 100m in length

ais_near_16 <- readLocalAIS(tracks_SS_16, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
   filter(vesselLength >= 100)  #remove vessels under 100m in length

ais_near_18 <- readLocalAIS(tracks_SS_18, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
   filter(vesselLength >= 100)  #remove vessels under 100m in length

ais_near_19 <- readLocalAIS(tracks_SS_19, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
   filter(vesselLength >= 100)  #remove vessels under 100m in length

ais_near_20 <- readLocalAIS(tracks_SS_20, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
   filter(vesselLength >= 100)  #remove vessels under 100m in length

ais_near_21 <- readLocalAIS(tracks_SS_21, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
   filter(vesselLength >= 100)  #remove vessels under 100m in length
#empty table because no large vessels interacted with this buoy

ais_near_22 <- readLocalAIS(tracks_SS_22, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
   filter(vesselLength >= 100)  #remove vessels under 100m in length

ais_near_23 <- readLocalAIS(tracks_SS_23, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
   filter(vesselLength >= 100)  #remove vessels under 100m in length
```

Match AIS points with buoy points
```{r}
gpsClose07 <- addAIS(tracks_SS_07, ais_near_07, interpType = 'close', interpTime = 120)
gpsClose08 <- addAIS(tracks_SS_08, ais_near_08, interpType = 'close', interpTime = 120)
gpsClose10 <- addAIS(tracks_SS_10, ais_near_10, interpType = 'close', interpTime = 120)
gpsClose12 <- addAIS(tracks_SS_12, ais_near_12, interpType = 'close', interpTime = 120)
gpsClose13 <- addAIS(tracks_SS_13, ais_near_13, interpType = 'close', interpTime = 120)
gpsClose14 <- addAIS(tracks_SS_14, ais_near_14, interpType = 'close', interpTime = 120)
gpsClose16 <- addAIS(tracks_SS_16, ais_near_16, interpType = 'close', interpTime = 120)
gpsClose18 <- addAIS(tracks_SS_18, ais_near_18, interpType = 'close', interpTime = 120)
gpsClose19 <- addAIS(tracks_SS_19, ais_near_19, interpType = 'close', interpTime = 120)
gpsClose20 <- addAIS(tracks_SS_20, ais_near_20, interpType = 'close', interpTime = 120)
gpsClose21 <- addAIS(tracks_SS_21, ais_near_21, interpType = 'close', interpTime = 120) #empty table
gpsClose22 <- addAIS(tracks_SS_22, ais_near_22, interpType = 'close', interpTime = 120)
gpsClose23 <- addAIS(tracks_SS_23, ais_near_23, interpType = 'close', interpTime = 120)
```

Create column 'shipsin10km' that displays number of ships associated with timestamp
Need to make a loop!
```{r}
# Viewing only ships within 10 km of buoys.
close2ships7 <- gpsClose07 %>% filter(shipDist < 10e3)

#Add column of ROUNDED UTC values to the lowest minute. Allows common field to bind
z <- as.POSIXct(close2ships7$UTC)
close2ships7$UTCrnd <- lubridate::floor_date(z, "minute")

#Bind rows with ship nearby
#Summarize unique ships and display number of nearby vessels
nearbyShp07 <- inner_join(tracks_ss_07, close2ships7, by = join_by("UTC" == "UTCrnd")) %>%
  select(UTC, UTC2, MMSI, vesselLength, vesselType, shipLat, shipLong, shipDist) %>%
  group_by(UTC2) %>%
  summarize(shipin10km = n_distinct(MMSI))

# ready to join to main df and add zeros to Null values
```
